# Data description

```{r Packages, echo = FALSE, message = FALSE}
source(here::here("scripts","setup.R"))
data <- read.csv(here("data", "Vehicle MPG - 1984 to 2023.csv"))
```

For this project, we selected a dataset focused on vehicle characteristics, available as a .csv file from data.world. You can access the dataset via the following link: [data.world](https://data.world/project-data-viz/vehicle-mpg-1984-to-2023-pdv-02-2023). It includes a total of 26 features describing 45,896 vehicle models from 141 brands released between 1984 and 2023. Below is a table providing an overview of the available features and their descriptions. You can find a deeper description of the data in [Annex -@sec-data_summary]. The analysis of the missing values will be dealt with in the data cleaning section.

### Description of the features

```{r echo=FALSE}
D1_1_SDG_table_info <- data.frame(
  Column1 = c("ID",
              "Model Year",
              "Make",
              "Model",
              "Estimated Annual Petroleum Consumption (Barrels)", 
              "Fuel Type 1",
              "City MPG (Fuel Type 1)",
              "Highway MPG (Fuel Type 1)",
              "Combined MPG (Fuel Type 1)",
              "Fuel Type 2",
              "City MPG (Fuel Type 2)",
              "Highway MPG (Fuel Type 2)",
              "Combined MPG (Fuel Type 2)",
              "Engine Cylinders",
              "Engine Displacement",
              "Drive",
              "Engine Description",
              "Transmission",
              "Vehicle Class",
              "Time to Charge EV (hours at 120v)",
              "Time to Charge EV (hours at 240v)",
              "Range (for EV)",
              "City Range (for EV - Fuel Type 1)",
              "City Range (for EV - Fuel Type 2)",
              "Hwy Range (for EV - Fuel Type 1)",
              "Hwy Range (for EV - Fuel Type 2)"),
  Column2 = c("Number corresponding to the precise combination of the features of the model", 
              "Year of the model of the car",
              "Brand of the car",
              "The model of the car",
              "Consumption in Petroleum Barrels",
              "First fuel energy source, only source if not an hybrid car", 
              "Consumption of the car in miles per gallon of fuel when driving in a city, for fuel type 1",
              "Consumption of the car in miles per gallon of fuel when driving on a highway, for fuel type 1",
              "Combined city and highway car consumption in miles per gallon, for fuel type 1",
              "Second energy source if hybrid car",
              "Consumption of the car in miles per gallon of fuel when driving in a city, for fuel type 2",
              "Consumption of the car in miles per gallon of fuel when driving on a highway, for fuel type 2",
              "Combined city and highway car consumption in miles per gallon, for fuel type 2",
              "Number of cylinders of the car",
              "Measure of the cylinder volume swept by all of the pistons of a piston engine, excluding the combustion chambers",
              "Type of powertrain distribution system that places rotational propulsion, such as rear-wheel,4-Wheel Drive,...",
              "Description of some features of the car, such as turbo engine, Stop-Start system, ...",
              "Manual/Automatic transmission, with number of gears and/or model of transmission",
              "Type of vehicle, such as Minivan, Trucks,...",
              "Number of hours required to fully charge an EV car at 120v",
              "Number of hours required to fully charge an EV car at 240v",
              "Maximum number of miles possible with a fully charged EV car",
              "Maximum number of miles possible with a fully charged EV car in a city",
              "Maximum number of miles possible while only using electricity with a fully charged hybrid car in a city",
              "Maximum number of miles possible with a fully charged EV car on a highway",
              "Maximum number of miles possible while only using electricity with a fully charged hybrid car on a highway"))

# Rename the columns
colnames(D1_1_SDG_table_info) <- c("Variable Name",
                                   "Explanation")

# Print the table with the variables information
D1_1_SDG_table_info %>%
  kable(escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped",
                                      "hover",
                                      "condensed",
                                      "responsive")) %>%
  column_spec(1, width = "35%")
```
