# Data description
* Description of the data file format (xlsx, csv, text, video, etc.) __DONE__
* The features or variables: type, units, the range (e.g. the time, numerical, in weeks from January 1, 2012 to December 31, 2015), their coding (numerical, the levels for categorical, etc.), etc. __TABLE-NTBF__
* The instances: customers, company, products, subjects, etc. __DONE__
* Missing data pattern: if there are missing data, if they are specific to some features, etc. __NTBD__
* Any modification to the initial data: aggregation, imputation in replacement of missing data, recoding of levels, etc. __NTBD__
* If only a subset was used, it should be mentioned and explained; e.g. inclusion criteria. Note that if inclusion criteria do not exist and the inclusion was an arbitrary choice, it should be stated as such. One should not try to invent unreal justifications. __NTBD__

```{r Packages, echo = FALSE, message = FALSE}
source(here::here("scripts","setup.R"))
data <- read.csv(here("data", "Vehicle MPG - 1984 to 2023.csv"))
```

For this project, we choose to go with a dataset on vehicle caracteristics. We found a `.csv` file on data.world. You can find the dataset using this link: [data.world](https://data.world/project-data-viz/vehicle-mpg-1984-to-2023-pdv-02-2023)
The dataset contains a total of 26 features (or caracteristics) for 45'896 models of vehicles released between 1984 and 2023. Find below a table that gives an overview of the features we have, its explanation and the type of values it is ((numerical, the levels for categorical, etc.).

### The features or variables: type, units,...

```{r echo=FALSE}
D1_1_SDG_table_info <- data.frame(
  Column1 = c("ID",
              "Model Year",
              "Make",
              "Model",
              "Estimated Annual Petroleum Consumption (Barrels)", 
              "Fuel Type 1",
              "City MPG (Fuel Type 1)",
              "Highway MPG (Fuel Type 1)",
              "Combined MPG (Fuel Type 1)",
              "Fuel Type 2",
              "City MPG (Fuel Type 2)",
              "Highway MPG (Fuel Type 2)",
              "Combined MPG (Fuel Type 2)",
              "Engine Cylinders",
              "Engine Displacement",
              "Drive",
              "Engine Description",
              "Transmission",
              "Vehicle Class",
              "Time to Charge EV (hours at 120v)",
              "Time to Charge EV (hours at 240v)",
              "Range (for EV)",
              "City Range (for EV - Fuel Type 1)",
              "City Range (for EV - Fuel Type 2)",
              "Hwy Range (for EV - Fuel Type 1)",
              "Hwy Range (for EV - Fuel Type 2)"),
  Column2 = c("Number corresponding to the precise combination of the features", 
              "Year of the model of the car",
              "The brand of the car, character",
              "The model of the car, character",
              "Consumption in Barrels, numerical",
              "Regular Gasoline, Diesel, ",
              "Score on each SDG except SDG 14 (16 variables)",
              "Population of the country",
              "d",
              "d",
              "d",
              "d",
              "d",
              "d",
              "d",
              "d",
              "d",
              "d",
              "d",
              "d",
              "d",
              "d",
              "d",
              "d",
              "d",
              "d"))

# Rename the columns
colnames(D1_1_SDG_table_info) <- c("Variable Name",
                                   "Explanation")

# Print the table with the variables information
D1_1_SDG_table_info %>%
  kable(escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped",
                                      "hover",
                                      "condensed",
                                      "responsive")) %>%
  column_spec(1, width = "35%")
```

## The instances: customers, company, products, subjects, etc.

In a basic instance, each row is concerning one car. We can find in order the ID of the car corresponding to a precise feature observation, then the features as seen in the table before. 

## Missing data pattern: if there are missing data, if they are specific to some features, etc.


## Any modification to the initial data: aggregation, imputation in replacement of missing data, recoding of levels, etc.


## If only a subset was used, it should be mentioned and explained; e.g. inclusion criteria. Note that if inclusion criteria do not exist and the inclusion was an arbitrary choice, it should be stated as such. One should not try to invent unreal justifications.


