---
format:
  html:
    code-fold: true
    fig-align: center
---

```{r Packages and data loading, echo = FALSE, message = FALSE, warning=FALSE}
source(here::here("scripts","setup.R"))
# data <- read.csv(here("data", "Vehicle_MPG_1984_to_2023_renamed.csv"), na = c(""))
#a voir pk fonctionne pas
data <- read_csv(here("data","data_initial_underscores.csv"))
```

# Exploratory Data Analysis

Now that we have presented the variables contained in the dataset, let's try to understand the data structure, characteristics and underlying patterns thanks to an exploratory data analysis.

### Dataset overview

Find below a preview of our raw dataset.

```{r, echo = FALSE, message = TRUE, warning = FALSE}
# Cleaned dataset
data <- data %>%
  complete(make)
datatable(data, options = list(pageLength = 5,
                                   class = "hover",
                                   searchHighlight = TRUE), rownames = FALSE)
```

### Columns description

Let's have a quick look at the characteristics of the columns. You will find more statistical details about it in the [Annex -@sec-data_column_details]. 

```{r, echo= FALSE}

# Get the number of rows and columns
num_rows <- nrow(data)
num_cols <- ncol(data)

# Get the frequency of column types
col_types <- sapply(data, class)
type_freq <- table(col_types)
char_count <- ifelse("character" %in% names(type_freq), type_freq["character"], 0)
num_count <- ifelse("numeric" %in% names(type_freq), type_freq["numeric"], 0)

# Create a summary data frame
summary_df <- data.frame(
  Name = "data",
  Number_of_rows = num_rows,
  Number_of_columns = num_cols,
  Character = char_count,
  Numeric = num_count,
  Group_variables = "None"
)

# Display the summary using kable
kable(summary_df, format = "html", table.attr = "style='width:50%;'") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```


The dataset that we are working with contains approximately 46'000 rows and 26 columns, each row representing a model from one of the 141 brands. From the data overview, we can see that most of our features are concerning the consumption of the cars. If we now check more in details in the annex, we notice that some variables contain a lot of missing and that the variable "Time.to.Charge.EV..hours.at.120v." is only containing 0s. We will handle these issues in the section "Data cleaning".

### Exploration of the distribution

Now let's explore the distribution of the numerical features.

```{r, fig.height= 5, fig.width=8, echo= FALSE}
#  melt.data <- melt(data)
# 
#  ggplot(data = melt.data, aes(x = value)) + 
# stat_density() + 
# facet_wrap(~variable, scales = "free")
plot_histogram(data)# Time.to.Charge.EV..hours.at.120v. not appearing because all observations = 0 
```

As the majority of models in our dataset are neither electric vehicles (EVs) nor hybrid cars and because of the nature of some column concerning only these two types of vehicles, the results are showing numerous zero values in several columns. This issue will be addressed during the data cleaning process. Additionally, certain features, such as "Engine Cylinders," are numerically discrete, as illustrated in the corresponding plot.

### Outliers Detection

In order to identify occurrences that deviate significantly for the rest of the observations, and in order to potentially improve the global quality of the data, we decided to analyse outliers thanks to boxplots. Here are the result on the numerical features of the dataset: 

```{r, fig.height=10, fig.width=12, echo= FALSE}
#tentative boxplots

data_long <- data %>%
  select_if(is.numeric) %>%
  pivot_longer(cols = c("ID", 
                        "model_year", 
                        "estimated_Annual_Petrolum_Consumption_Barrels", "City_MPG_Fuel_Type_1",
                        "highway_mpg_fuel_type_1", 
                        "combined_MPG_Fuel_Type_1", 
                        "City_MPG_Fuel_Type_2", 
                        "highway_mpg_fuel_type_2", 
                        "combined_MPG_Fuel_Type_2", 
                        "time_to_Charge_EV_hours_at_120v_", 
                        "charge_time_240v", 
                        "range_for_EV", 
                        "range_ev_city_fuel_type_1", 
                        "range_ev_city_fuel_type_2", 
                        "range_ev_highway_fuel_type_1", 
                        "range_ev_highway_fuel_type_2"), names_to = "variable", values_to = "value")

ggplot(data_long, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(outlier.size = 0.5) +  # Make outlier points smaller
  facet_wrap(~ variable, scales = "free_y") +  # Each variable gets its own y-axis
  theme_minimal() +
  theme(legend.position = "none",  # Hide the legend
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 0),strip.text = element_text(size = 7)) +  # Rotate x-axis labels
  labs(title = "Outlier Detection Boxplots", x = "", y = "Values")

```

Most of our boxplots are showing extreme values. Again, this is due to the small amount of EV and hybrid cars in our dataset compared to the rest of the models and due to the nature of some features, concerning only those type of vehicles.

```{r, echo=FALSE}
#Now 
# plot_correlation(data) #drop time charge EV 120V
# create_report(data)
#nb cars per brand
```

### Number of models per make

Now let's check how many models per make we have in our dataset. In order to have a clear plot, we decided to keep only the top 20 brands among all the make in our dataset to create the plot. All the remaining makes are accessible on in [Annex -@sec-data_n_make]. 

```{r, echo= FALSE}
#Number of occurences/model per make 
# nb_model_per_make <- data %>%
#   group_by(make, model) %>%
#   summarise(Number = n(), .groups = 'drop') %>%
#   group_by(make) %>%
#   summarise(nb_model_per_make = n(), .groups = 'drop') %>%
#   arrange(desc(nb_model_per_make))

nb_model_per_make <- data %>%
  group_by(make) %>%
  summarise(Number = n(), .groups = 'drop')

#makes with only 1 model 
only_one_model_cars <- nb_model_per_make %>%
  filter(Number == 1 )

#table globale
# datatable(nb_model_per_make,
#           rownames = FALSE,
#           options = list(pageLength = 5,
#                          class = "hover",
#                          searchHighlight = TRUE))
```

```{r, echo= FALSE}
# Option to limit to top 20 makes for better readability
# top_n_makes <- nb_model_per_make %>% top_n(20, nb_model_per_make)

top_n_makes <- nb_model_per_make %>%
  arrange(desc(Number)) %>%  # Sort in descending order by the column 'count'
  slice_max(Number, n = 20)

#plot
ggplot(top_n_makes, aes(x = reorder(make, Number), y = Number)) +
  geom_bar(stat = "identity", color = "black", fill = "grey", show.legend = FALSE) +
  labs(title = "Models per Make (Top 20)",
       x = "Make",
       y = "Number of Models") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(hjust = 1, size = 10),
        plot.title = element_text(size = 14)) +
  coord_flip()  # Flip coordinates for better readability
```

On the 141 brands, we notice that only 13 brands have more than 1000 models in the dataset. Among these, only one of them (Chevrolet) have more than 4000 models presents. In addition, we saw that many car brands have 1 observation, which would lead to class unbalanced and would lead to bias toward majority classes. We will address this issue later.

```{r, fig.height=7, echo= FALSE}
#table only 1 model
# datatable(only_one_model_cars,
#           rownames = FALSE,
#           options = list(pageLength = 5,
#                          class = "hover",
#                          searchHighlight = TRUE))
# 
# ggplot(only_one_model_cars, aes(x = make, y = Number)) +
#   geom_bar(stat = "identity", color = "black", fill = "grey", show.legend = FALSE) +
#   labs(title = "Makes with only 1 model in the dataset",
#        x = "Make",
#        y = "Number of Models") +
#   theme_minimal() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
#         axis.text.y = element_text(hjust = 1, size = 10),
#         plot.title = element_text(size = 14)) +
#   coord_flip()  # Flip coordinates for better readability

```

